Exploring dimensions of biodiversity in Japanese ferns

### 2021-07-26 ms/manuscript.Rmd
Update MS
add maps of genera, barplot of % apo

```{r setup, include = FALSE}
# Run these lines if rendering outside of targets plan
# library(targets)
# library(tarchetypes)
# source(here::here("R/packages.R"))
# source(here::here("R/functions.R"))

# Load libraries used for rendering Rmds
# library(officedown)
# library(officer)
library(flextable)
library(kableExtra)
library(ftExtra)
library(patchwork)

@@

```{r load-data}
# Load objects from targets workflow
tar_load(c(
  "occ_point_data_ferns",
  "occ_point_data_ferns_unfiltered",
  "biodiv_ferns_spatial",
  "k_taxonomy",
  "k_phylogeny",
  "signif_cells_protected_area",
  "protected_areas",
  "japan_shp",
  "phy_sig_results",
  "binary_sig_results",
  "traits_summary",
  "fern_traits",
  "japan_pterido_tree",
  "japan_fern_tree",
  "ppgi",
  "occ_point_data_ferns",
  "comm_ferns_full",
  "comm_ferns",
  "japan_points_raw",
  # "plastome_calibration_dates",
  # FIXME: remove until phylo analysis is working "plastome_tree",
  "green_list",
  "model_params",
  "lrt_results",
  "aic_env_repro",
  "model_fits",
  "biodiv_ferns_cent"
))
```

@@

^2^University and Jepson Herbaria, University of California, Berkeley, CA, USA

@@

**Methods** We used a list of >300,000 fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa. 
We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and modeled the distribution of biodiversity in response to environmental factors and reproductive mode. 
We conducted categorical analysis of neo- and paleo-endemism to identify areas of significant phylogenetic endemism. 
We used hierarchical clustering of taxonomic and phylogenetic distances to analyze bioregions. 
We compared the overlap of significantly diverse grid-cells with existing conservation zones to assess conservation status.

**Results** We observe three major patterns in the spatial distribution of biodiversity: that of taxonomic richness, which is highest at mid-latitudes, that of phylogenetic and functional diversity and phylogenetic endemism, which are highest in small southern islands, and that of relative phylogenetic and functional diversity, which are high at both high and low latitudes, and low at mid-latitudes. 
Grid-cells were grouped in to three (phylogenetic) or four (taxonomic) major bioregions. 
Conservation status (% protected area) is higher than the overall rate for grid-cells with significantly high biodiversity.

@@

Characterizing the spatial distribution of biodiversity is a major goal of evolutionary biology with two equally important aspects: to understand the processes generating it, and to conserve it. 
The vast majority of studies seeking to understand the spatial distribution of biodiversity focus on species richness. 
Hotspots, or areas of exceptionally high richness, have received particular attention, resulting in a widely-recognized set of 34 terrestrial hotspots [@Myers2000] and motivating conservation strategies to preserve them [@Margules2000; @Brooks2006]. 
However, richness alone cannot provide a complete picture of biodiversity. 
All organisms are related by descent from a common ancestor, and these evolutionary relationships must be taken into account to obtain a full understanding of biodiversity.

Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because data (DNA sequences) and analytic tools have only become available relatively recently. 
These tools and datasets now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity and phylogenetic endemism. 
Better ways to characterize biological regions ("bioregions") are also available, rather than relying on ad-hoc characterizations [e.g., @White2019; @Daru2020a]. 
Furthermore, incorporating these two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide particularly powerful insights into the processes structuring biodiversity and suggest conservation priorities.

However, a comprehensive understanding of the relationships between richness and other metrics requires densely sampled data, and such datasets are rare on the regional (country) scale. 
The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled. 
One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for 221/721 taxa [70%]), which is likely to affect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].

@@

We used a list of all non-hybrid fern specimens housed at herbaria in Japan. 
We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/). 
Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")`). 
The original list comprises `r nrow(occ_point_data_ferns_unfiltered) %>% scales::number(big.mark = ",")` specimens representing `r occ_point_data_ferns_unfiltered %>% pull(taxon) %>% n_distinct()` taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated). 
We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html).

A necessary step in any analysis of biodiversity is to set the grain size used to define communities. 
Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. 
We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of a grid-cell) for defining communities (co-occurring species). 
At each grain size, species occurrences were converted to a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell). 
We calculated sampling redundancy (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling. 
Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`). 
We filtered out any cells with redundancy < 0.1 and used this for subsequent analyses.

@@

We used traits originally compiled for identification of ferns and lycophytes of Japan as follows [@Ebihara2016b; @Ebihara2017]. 
Continuous traits were measured on 10 randomly selected specimens per species. 
Fertile (`r ie`, spore-bearing) and sterile fronds were measured separately, then averaged. 
Qualitative traits were scored by observing voucher specimens. 
All qualitative traits were scored in binary format; `r eg`, a single trait with three states was formatted as three binary traits. 
From the original trait list, we selected only traits with putative ecological function (`r table("traits-used")`), and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.

@@

We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")`, and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`). 
We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
We calculated area as rolling mean area (km^2^) across 1 degree latitudinal bands.
This approach reflects the idea that the amount of area with a given climate regime enclosing each grid cell may influence species richness of that cell [@Fahrig2013], and that simply using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by.
The R package `r pack("sf")` was used for GIS analyses.

@@

It is possible that endemism may be affected by species that are actually wide-ranging, but only have a small range in Japan (the "border effect"). 
This is especially expected for southern, sub-tropical islands, which harbor many tropical species with small ranges in Japan but wider ranges outside of the country. 
To account for this, we also calculated all endemism-related metrics using a dataset including only species endemic to Japan.

@@

We postulated that % apomictic taxa might explain the distribution of genetic diversity better than temperature, so we also constructed a limited set of models to test this hypothesis ("reproductive models"). 
These models included biodiversity metrics with a genetic component (SES of PD and SES of RPD) as dependent on % apomictic taxa, precipitation, and precipitation seasonality, with random spatial effects as described above. 
We compared the fit between environmental models and reproductive models for SES of PD and SES of RPD using conditional Akaike Information Criterion (cAIC) [@Courtiol2017; @Vaida2005].

@@

Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_biogeo. 
A Docker image to run the code is available at https://hub.docker.com/repository/docker/joelnitta/japan_ferns_biogeo.
Data are openly available from Figshare at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.

@@

Taxonomic richness is lowest on the northernmost island of Hokkaido, then increases with decreasing latitude until reaching a maximum in southern Kyushu, then decreases again in smaller islands further south (`r figure("raw-diversity")`A). 
Observed values of PD and FD were highly correlated with richness as expected (`r figure("raw-diversity")`B, C), showing an initial steep slope that gradually starts to level off at higher richness (`r s_figure("raw-biplot")`A, C). 
Observed values of RPD and RFD were also closely related to richness, showing high variance at low species richness, and low variance at high species richness (`r s_figure("raw-biplot")`B, D).

@@

Nearly all grid-cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere (`r figure("endemism")`). 
The vast majority of endemic grid-cells are mixed. 
A cluster of paleoendemic cells was observed in the Ryukyu Islands. 
Small clusters of significantly endemic cells were also observed northwest of Tokyo (Nagano prefecture), and in central Hokkaido (Daisetsuzan mountains). 
When we ran CANAPE with a reduced dataset including only taxa endemic to Japan to account for the border effect, there were fewer significant cells, but patterns were otherwise similar to the full dataset (`r s_figure("endemism-restricted")`).

@@

Here, we provide the first insights into the combined taxonomic, phylogenetic, and functional diversity of the ferns of Japan. 
Our integrative approach reveals distinct patterns of biodiversity in different parts of the archipelago and across biodiversity metrics, and suggest reproductive mode as an important driver in some cases. 
These clear differences are reflected in the bioregionalization results, which delimit three to four major bioregions. 
Finally, we consider the current conservation status of ferns in Japan and the implications of our findings for future conservation activity.

@@

The latitudinal hump-shaped pattern of species richness observed in ferns here (`r figure("raw-diversity")`A) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015] and modeled distributions of woody plants [@Fukaya2020]. 
Our modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`A).
Such a hump-shaped pattern is often observed in ferns on mountains, particularly in tropical or sub-tropical areas [@Kessler2011a]. 
Hypotheses to explain such a pattern include pure geometrical constraints as well as biological hypothesis positing maximally favorable environmental conditions at mid-elevations [@Colwell2004a]. 
Another reason for the greater richness in southern Honshu compared to smaller southern islands is that southern Honshu harbors high elevational variation that may support a wide range of taxa that is not present in the southern islands.

@@

Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half on Honshu, particularly Kyushu (`r figure("rand-diversity")`B). 
The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* and *Botrychium*. The preponderance of short branches in southern Honshu may be due to the radiation of certain species-rich genera such as *Dryopteris* and *Deparia*.

Furthermore, we identified % apomictic taxa as a strong driver of SES of PD (`r figure("modelfit")`G); indeed, *Dryopteris* and *Deparia* are two such genera with high numbers of apomictic taxa. 
This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower genetic diversity, and is the first time to our knowledge that this has been demonstrated in ferns. 
While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD. 
Ultimately, it is likely that temperature, in combination with other environmental conditions, is driving the predominance of apomictic taxa: @Tanaka2014 also found that the proportion of apomictic taxa reaches a maximum at ca. 34.8 °N latitude and increases with temperature in Japan (their study was limited to Honshu and did not include the southern islands). 
@Tanaka2014 assert that one limiting factor affecting establishment of apomictic ferns at high latitudes or cold climates is the large cell size of apomicts, which are typically polyploid; as fern gametophytes are only a single cell-layer in thickness with no cuticle or stomata, it is expected they would be particularly susceptible to freezing damage induced by larger cells with higher water content; however, no experimental studies have tested these hypotheses.

@@

The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group cells with similar biodiversity values (`r figure("div-by-tax-region")`). 
Interestingly, the bioregions identified here correspond roughly to major forest types previously categorized on an *ad-hoc* basis; Cluster 1 corresponds to the cool or subartic deciduous type, Cluster 2 to the warm evergreen type, and Clusters 3 and 4 to the subtropical rainy type [@Shimizu2014]. 
This suggests that ferns are useful as possible bioindicators in Japan. 
Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from the rest of Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions. 
The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits clusters 2 and 3, there is some overlap between them on the Amami-Oshima Isl (`r figure("bioregions")`). 
The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind.
Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic Cluster 3 (mostly the Ryukyu islands, wetter) from Cluster 4 (mostly the Ogasawara islands, drier) (`r s_figure("bioregions-env")`).

>>>>

```{r manual, include = FALSE, evaluate = FALSE}
# Run these lines if rendering outside of targets plan
library(targets)
library(tarchetypes)
source(here::here("R/packages.R"))
source(here::here("R/functions.R"))
```

```{r setup, include = FALSE}
# Load libraries used for rendering Rmds
# library(officedown)
# library(officer)
library(flextable)
library(kableExtra)
library(ftExtra)
library(patchwork)

@@

```{r load-data}
# Load objects from targets workflow
tar_load(c(
  "repro_data",
  "occ_point_data_ferns",
  "occ_point_data_ferns_unfiltered",
  "biodiv_ferns_spatial",
  "k_taxonomy",
  "k_phylogeny",
  "signif_cells_protected_area",
  "protected_areas",
  "japan_shp",
  "phy_sig_results",
  "binary_sig_results",
  "traits_summary",
  "fern_traits",
  "japan_pterido_tree",
  "japan_fern_tree",
  "ppgi",
  "occ_point_data_ferns",
  "comm_ferns_full",
  "comm_ferns",
  "japan_points_raw",
  # "plastome_calibration_dates",
  # FIXME: remove until phylo analysis is working "plastome_tree",
  "green_list",
  "model_params",
  "lrt_results",
  "aic_env_repro",
  "model_fits",
  "biodiv_ferns_cent"
))
```

@@

^2^University and Jepson Herbaria, and Department of Integrative Biology, University of California, Berkeley, CA, USA

@@

**Methods** We used data from >300,000 fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa. 
We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and modeled the distribution of biodiversity in response to environmental factors and reproductive mode. 
We conducted categorical analysis of neo- and paleo-endemism to identify areas of significant phylogenetic endemism. 
We used hierarchical clustering of taxonomic and phylogenetic distances to analyze bioregions. 
We compared the overlap of significantly diverse grid-cells with existing conservation zones to assess conservation status.

**Results** We observe three major patterns in the spatial distribution of biodiversity: that of taxonomic richness, which is highest at mid-latitudes, that of phylogenetic and functional diversity and phylogenetic endemism, which are highest in small southern islands, and that of relative phylogenetic and functional diversity, which are high at both high and low latitudes, and low at mid-latitudes. 
Grid-cells were grouped in to three (phylogenetic) or four (taxonomic) major bioregions. 
Temperature and apomixis were identified as likely drivers of patterns of taxonomic and phylogenetic diversity.
Conservation status (% protected area) is higher than the overall rate for grid-cells with significantly high biodiversity.

@@

```{r intro-stats}
# tally number of non-hybrid pteridophytes with known repro. mode
repro_count <-
repro_data %>%
  count(reproductive_mode) %>%
  mutate(mode = if_else(reproductive_mode == "unknown", "unknown", "known")) %>%
  group_by(mode) %>%
  summarize(total = sum(n)) %>%
  split(.$mode)

repro_count$known$total

repro_count$known$percent <- (repro_count$known$total/(repro_count$unknown$total + repro_count$known$total)) %>% percent(accuracy = 0.1)
```

Characterizing the spatial distribution of biodiversity is a major goal of evolutionary biology with two equally important aspects: to understand the processes generating it, and to conserve it. 
The vast majority of studies seeking to understand the spatial distribution of biodiversity focus on species richness. 
Hotspots, or areas of exceptionally high richness, have received particular attention, resulting in a widely-recognized set of 34 terrestrial hotspots [@Myers2000] and motivating conservation strategies to preserve them [@Margules2000; @Brooks2006]. 
However, richness alone cannot provide a complete picture of biodiversity. 
All organisms are related to a greater or lesser degree by descent from a common ancestor, and these evolutionary relationships must be taken into account to obtain a full understanding of the biodiversity present in an area.

Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because data (DNA sequences), extensive spatial data, and analytic tools have only become available relatively recently. 
These tools and datasets now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity and phylogenetic endemism. 
Better ways to characterize biological regions ("bioregions") are also available, rather than relying on ad-hoc characterizations [e.g., @Laffan2016; @White2019; @Daru2020a].
Furthermore, incorporating these two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide particularly powerful insights into the processes structuring biodiversity and suggest conservation priorities.

However, a comprehensive understanding of the relationships between richness and other metrics requires densely sampled data, and such datasets are rare on the regional (country) scale. 
The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled. 
One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa excluding hybrids [`r repro_count$known$percent`]), which is likely to affect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].

@@

We used a list of all non-hybrid fern specimens housed at herbaria in Japan. 
We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/). 
Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")`). 
The original list comprises `r nrow(occ_point_data_ferns_unfiltered) %>% scales::number(big.mark = ",")` specimens representing `r occ_point_data_ferns_unfiltered %>% pull(taxon) %>% n_distinct()` terminal taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated). 
We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html).

A necessary step in any analysis of biodiversity is to set the grain size used to accurately define communities. 
Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. 
We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of a grid-cell) for defining communities (co-occurring species). 
At each grain size, species occurrences were converted to a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell). 
We calculated sampling redundancy (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling. 
Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`). 
We filtered out any cells with redundancy < 0.1 and used this size for subsequent analyses.

@@

We used traits originally compiled for identification of ferns and lycophytes of Japan following Ebihara [-@Ebihara2016b; -@Ebihara2017]. 
Continuous traits were measured on 10 randomly selected specimens per species. 
Fertile (`r ie`, spore-bearing) and sterile fronds were measured separately, then averaged. 
Qualitative traits were scored by observing voucher specimens. 
All qualitative traits were scored in binary format; `r eg`, a single trait with three states was formatted as three binary traits. 
From the original trait list, we selected only traits with putative ecological function (`r table("traits-used")`), and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.

@@

We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")`, and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`). 
We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
We calculated area as rolling mean area (km^2^) across 1 degree latitudinal bands.
This approach reflects the idea that the amount of area with a given climate regime enclosing each grid cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by.
The R package `r pack("sf")` was used for GIS analyses.

@@

It is possible that endemism may be affected by species that are actually wide-ranging, but only have a small range in Japan (a "border effect"). 
This is especially expected for southern, sub-tropical islands, which harbor many tropical species with small ranges in Japan but wider ranges outside of the country. 
To partially account for this, we also calculated all endemism-related metrics using a dataset including only species endemic to Japan.

@@

We postulated that % apomictic taxa might explain the distribution of phylogenetic diversity better than temperature, so we also constructed a limited set of models to test this hypothesis ("reproductive models"). 
These models included biodiversity metrics with a phylogenetic component (SES of PD and SES of RPD) as dependent on % apomictic taxa, precipitation, and precipitation seasonality, with random spatial effects as described above. 
We compared the fit between environmental models and reproductive models for SES of PD and SES of RPD using conditional Akaike Information Criterion (cAIC) [@Courtiol2017; @Vaida2005].

@@

Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_biogeo. 
A Docker image to run the code is available at https://hub.docker.com/repository/docker/joelnitta/japan_ferns_biogeo.
Data files, including the DNA alignment, phylogenetic tree, and community data matrix, are openly available from Figshare at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.

@@

Taxonomic richness is lowest on the northernmost island of Hokkaido, then increases with decreasing latitude until reaching a maximum in southern Kyushu, then decreases again in smaller islands further south (`r figure("raw-diversity")`A). 
Observed values of PD and FD were highly correlated with richness as expected (`r figure("raw-diversity")`B, C), showing an initial steep slope that gradually starts to level off at higher richness (`r s_figure("raw-biplot")`A, C). 
Observed values of RPD and RFD were not linearly related to richness, however, they showed high variance at low species richness, and low variance at high species richness (`r s_figure("raw-biplot")`B, D).

@@

Nearly all grid-cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere (`r figure("endemism")`). 
The vast majority of grid-cells with significant endemism are mixed.
A concentration of paleoendemism was observed in the Ryukyu Islands. 
Small clusters of significantly endemic cells were also observed northwest of Tokyo (Nagano prefecture), and in central Hokkaido (`r figure("endemism")`). 
When we ran CANAPE with a reduced dataset including only taxa absolutely endemic to Japan to account for the border effect, there were fewer significant cells, but patterns were otherwise similar to the full dataset (`r s_figure("endemism-restricted")`).

@@

@@

The latitudinal hump-shaped pattern of species richness observed in ferns here (`r figure("raw-diversity")`A) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants [@Fukaya2020]. 
Our spatial modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`A).
Previous studies have identified temperature, water availability [@Bickford2006; @Qian2012], or a combination of these [@Nagalingum2015; @Link-Perez2018a; @Qian2021] to be important drivers of spatial fern diversity.
A hump-shaped pattern of richness along elevation (maximum richness at intermediate elevations) has been frequently observed in ferns on mountains in tropical or sub-tropical areas [@Kessler2011a]. 
Hypotheses to explain such a pattern include purely geometrical constraints as well as biological hypothesis positing maximally favorable environmental conditions at mid-elevations [@Colwell2004a]. 
Another reason for the greater richness in southern Honshu compared to smaller southern islands is that southern Honshu harbors high elevational variation that may support a wider range of taxa than is present in the southern islands.

@@

Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half on Honshu, particularly Kyushu (`r figure("rand-diversity")`B). 
The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`A). The preponderance of short branches in southern Honshu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`B) and *Dryopteris* (`r s_figure("genus-map")`D).

Furthermore, we identified % apomictic taxa as a strong driver of SES of PD (`r figure("modelfit")`G).
Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in southern Honshu (`r s_figure("genus-map")`).
This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower genetic diversity, and is the first time to our knowledge that this has been demonstrated in ferns. 
While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD (`r s_table("aic")`). 
Ultimately, it is likely that temperature, in combination with other environmental conditions, is driving the predominance of apomictic taxa: @Tanaka2014 also found that the proportion of apomictic taxa reaches a maximum at ca. 34.8 °N latitude and increases with temperature in Japan (their study was limited to Honshu and did not include the southern islands). 
@Tanaka2014 assert that one limiting factor affecting establishment of apomictic ferns at high latitudes or cold climates is the large cell size of apomicts, which are typically polyploid. 
As fern gametophytes are only a single cell-layer in thickness with no cuticle or stomata, it is expected they would be particularly susceptible to freezing damage induced by larger cells with higher water content; however, no experimental studies have tested these hypotheses.

@@

The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group cells with similar biodiversity values (`r figure("div-by-tax-region")`). 
Interestingly, the bioregions identified here correspond roughly to major forest types previously categorized on an *ad-hoc* basis; Cluster 1 corresponds to the cool or subartic deciduous type, Cluster 2 to the warm evergreen type, and Clusters 3 and 4 to the subtropical rainy type [@Shimizu2014]. 
This suggests that ferns are useful as possible bioindicators in Japan. 
Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from the rest of Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions. 
The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits clusters 2 and 3, there is some overlap between them on the Amami-Oshima Isl. (`r figure("bioregions")`). 
The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind.
Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic Cluster 3 (mostly the Ryukyu islands; wetter) from Cluster 4 (mostly the Ogasawara islands; drier) (`r s_figure("bioregions-env")`).
